Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

In Harmony, \texttt{atomic} statements allow you to \emph{implement} your own
synchronization primitives like test-and-set.  Other common examples
include compare-and swap and fetch-and-add.  \texttt{atomic} statements
are not intended to \emph{replace} locks or other synchonization primitives.
When solving synchronization problems you should not directly use
\emph{atomic} statements but use the synchronization primitives that are available
to you.  But if you want to design a new synchronization primitive, then
use \textbf{atomic} by all means.

@@

The producer/consumer pattern is common.  Processes may be arranged
in \emph{pipelines},
\index{pipeline}%
where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers executing them in parallel.
Or, in the client/server model,
\index{client/server model}%
some process may act as a \emph{server} that clients can send requests to
and receive responses from.

In general, we allow multiple producers and multiple consumers
all sharing the same bounded buffer.
\autoref{fig:pc} gives a high-level description.  The
two numbers in a state specify
the number of items that the producers still can produce and
the number of items that the consumers have consumed.  In this case
there are three items to produce initially.

@@

\autoref{fig:bbtest} invokes the code of \autoref{fig:boundedbuffer}.
We can run this through Harmony, but what would it test?  There are no
assert statements in the code.
We can split the functionality of this code into two pieces.  One is
the synchronization part: producers should never run more than \texttt{NSLOTS}
ahead of the consumers, but they should be able to produce up to \texttt{NSLOTS}
items even if there are no consumers.  The other part is the data: we want
to make sure that the items are sent through the buffer in order and without
loss.  We will first focus on the synchronization.

>>>>

In Harmony, \texttt{atomic} statements allow you to \emph{implement} your own
synchronization primitives like test-and-set.  Other common examples
include compare-and swap and fetch-and-add.  \texttt{atomic} statements
are not intended to \emph{replace} locks or other synchonization primitives.
When solving synchronization problems you should not directly use
\emph{atomic} statements but use the synchronization primitives that are available
to you.  But if you want to design a new synchronization primitive, then
use \textbf{atomic} by all means.
You can also use \textbf{atomic} statements in your test code.
In fact, \texttt{assert} statements are executed atomically.

@@

The producer/consumer pattern is common.  Processes may be arranged
in \emph{pipelines},
\index{pipeline}%
where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers consuming and executing them in parallel.
Or, in the client/server model,
\index{client/server model}%
some process may act as a \emph{server} that clients can send requests to
and receive responses from.  In that case, there is a bounded buffer
for each client/server pair. Clients produce requests and
consume responses, while the server consumes requests and produces responses.

In general, we allow multiple producers and multiple consumers
all sharing the same bounded buffer.
\autoref{fig:pc} gives a high-level description.  The
two numbers in a state specify
the number of items that the producers still can produce and
the number of items that the consumers have consumed.  In this
description there are three items to produce initially.

@@

\autoref{fig:bbtest} invokes the code of \autoref{fig:boundedbuffer}.
We can run this through Harmony, but what would it test?  There are
no assert statements in the code.  We can split the functionality
of this code into two and test each separately.  One is the
synchronization part: producers should never run more than
\texttt{NSLOTS} ahead of the consumers, but they should be able to
produce up to \texttt{NSLOTS} items even if there are no consumers.
The other part is the data: we want to make sure that the items are
sent through the buffer in order and without loss.  We will first
focus on the synchronization.
